Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models

### 2020-08-16 implementation/implementation.tex
New fix

    \caption[Factors and adaptive distance experiments]%
    {Factors and adaptive distance experiments. Left: Total required samples. Right: acceptance rates in each generation}
    \label{fig:factor}

@@

    \caption[CAPTION]%
    {DESCRIBE}
    \label{fig:factor_sim}

@@

    \caption[Data size and prior distribution range experiment]%
    {Data size and prior distribution range experiment. (a) Total required samples, (b) Epsilon values under median epsilon schedule}
    \label{fig:size}

@@

    \caption{Simulated data from the last population of model 3, 4 and 5}
    \label{fig:resultCurve345}

@@

When we are trying more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}). In this case, the discrepancy of simulated data and observed data comes from the last 2 data point (72 hpl and 120 hpl). MORE

@@

    \caption[Simulated trajectory from the last population of model 5, with factors applied]{Simulated trajectory from the last population of model 5, with factors applied. MORE DETAIL}
    \label{fig:overfit}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model345cmp.png}}
    \end{center}

    \caption[ABC SMC based model comparison of model 3, 4 and 5]{ABC SMC based model comparison of model 3, 4 and 5. MORE DETAIL}
    \label{fig:model345cmp}
\end{figure}

@@

@@

For the best model (model 5), the inferred parameter posterior distribution of model 5 is shown in Figure \ref{fig:model5_para}, the estimated mean values is shown in Table \ref{table:estimated5}. The (pair-wise) joint density distribution is shown in Figure \ref{fig:model5_para_joint}. SOME DETAIL.

@@

    \caption[Estimated posterior distribution of parameters in model 5]%
    {Estimated posterior distribution of parameters in model 5. Shaded range indicates 25\%--75\% quantile of the population}
    \label{fig:model5_para}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c c c|}
        \hline
        Parameter            & Mean value & Inter-quartile interval \\[0.5ex]
        \hline\hline
        $\lambda_N$          & 13.5       & [12.9, 14.2]            \\
        $a$                  & 0.0175     & [0.0145, 0.0207]        \\
        $\kappa_{N\beta}$    & 0.176      & [0.109, 0.234]          \\
        $\mu_N$              & 0.0239     & [0.0113, 0.0332]        \\
        $\nu_{N\Phi}$        & 0.125      & [0.115, 0.134]          \\
        \hline
        $\kappa_{\Phi\beta}$ & 0.123      & [0.115, 0.131]          \\
        $\mu_\Phi$           & 0.0234     & [0.0199, 0.0267]        \\
        \hline
        $s_{\beta N}$        & 2.38       & [2.02, 2.73]            \\
        $\mu_\beta$          & 2.67       & [2.24, 3.10]            \\
        \hline
        $s_{\alpha\Phi}$     & 0.574      & [0.295, 0.786]          \\
        $\mu_\alpha$         & 8.91       & [6.07, 11.1]            \\
        $f_{\beta\alpha}$    & 0.577      & [0.361, 0.738]          \\
        \hline
    \end{tabular}
    \caption[Inferred parameter values of model 5]
    {Inferred parameter values of model 5. MORE DETAIL(SUPER maxt-5)}
    \label{table:estimated5}
\end{table}

@@

    \caption[PCA results of the last population of model 5]{PCA results of the last population of model 5. MORE DESCRIPTION}
    \label{fig:pc}
\end{figure}

@@

    \caption[TODO]{PCA results of the last population of model 5. MORE DESCRIPTION}
    \label{fig:pc_pie}
\end{figure}

>>>>

    \caption[Factors and adaptive distance experiments]%
    {Factors and adaptive distance experiments. Left: total required samples. Different color represents different generations (bottom totop: population 1 to population 20). Right: acceptance rates in each generation. }
    \label{fig:factor}

@@

    \caption[Simulated trajectories ]%
    {Simulated trajectories from the last population in the factor and adaptive distance experiment. 500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from. 25:75 factor showed very similar curve as the range factor.  
    }
    \label{fig:factor_sim}

@@

    \caption[Data size and prior distribution range experiment]%
    {Data size and prior distribution range experiment. Left: total required samples. Right: epsilon values under median epsilon schedule}
    \label{fig:size}

@@

    \caption{Simulated data from the last population of model 3, 4 and 5. Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
    \label{fig:resultCurve345}

@@

When we are trying more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}). In this case, the discrepancy of simulated data and observed data mainly comes from the last 2 data point (72 hpl and 120 hpl) of macrophage. As additional importance was assigned the the first few data points of each trajectory, the inference algorithm tended to preferentially fit first data points and thus relative large discrepancy at the tails were acceptable.

@@

    \caption[Simulated trajectory from the 30th population of model 5, with 25:75 factor applied]{Simulated trajectory from the 30th population of model 5, with 25:75 factor applied. Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small).}
    \label{fig:overfit}
\end{figure}

@@

\begin{figure}[H]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model345cmp.png}}
    \end{center}

    \caption[Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison]{Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison. Model 5 also wins in further populations (population 25 to 30)}
    \label{fig:model345cmp}
\end{figure}

@@

For the best model (model 5), the inferred parameter posterior distribution of model 5 is shown in Figure \ref{fig:model5_para}, the estimated mean values is shown in Table \ref{table:estimated5}. The pair-wise joint density distribution is shown in Figure \ref{fig:model5_para_joint}. Most of inferred posteriors show a bell-shaped density distribution, indicating that they are gaussian distributed. A thin and concentrated shape indicate a well inferred parameter, e.g. $\mu{\Phi}$ and $\kappa_{\Phi\beta}$. Some distributions, however, are skewed to zero, e.g. $\mu_N$ ans $s_{\alpha\Phi}$.

By looking at the joint density plot (Figure \ref{fig:model5_para_joint}), some correlated patterns are observed. For example, an obvious linear relationship on the distributions of $\mu_\beta$ and $s_{\beta\Phi}$. This leads us to consider that reducing the number of parameters in the model is possible. Highly correlated parameters can be reduced, e.g. $s_{\beta\Phi}$ can be expressed as $A\mu_\beta+B$ where $A$ and $B$ can be calculated from the joint distribution by fitting a linear regression.

@@

    \caption[Inferred posterior distribution of parameters in model 5]%
    {Inferred posterior distribution of parameters in model 5. Shaded range indicates 25\%--75\% quantile of the population}
    \label{fig:model5_para}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c c c|}
        \hline
        Parameter            & Mean value & Inter-quartile interval \\[0.5ex]
        \hline\hline
        $\lambda_N$          & 13.5       & [12.9, 14.2]            \\
        $a$                  & 0.0175     & [0.0145, 0.0207]        \\
        $\kappa_{N\beta}$    & 0.176      & [0.109, 0.234]          \\
        $\mu_N$              & 0.0239     & [0.0113, 0.0332]        \\
        $\nu_{N\Phi}$        & 0.125      & [0.115, 0.134]          \\
        \hline
        $\kappa_{\Phi\beta}$ & 0.123      & [0.115, 0.131]          \\
        $\mu_\Phi$           & 0.0234     & [0.0199, 0.0267]        \\
        \hline
        $s_{\beta N}$        & 2.38       & [2.02, 2.73]            \\
        $\mu_\beta$          & 2.67       & [2.24, 3.10]            \\
        \hline
        $s_{\alpha\Phi}$     & 0.574      & [0.295, 0.786]          \\
        $\mu_\alpha$         & 8.91       & [6.07, 11.1]            \\
        $f_{\beta\alpha}$    & 0.577      & [0.361, 0.738]          \\
        \hline
    \end{tabular}
    \caption[Inferred parameter values of model 5]
    {Inferred parameter values of model 5}
    \label{table:estimated5}
\end{table}

@@

    \caption[Explained total variance in original parameters by each PC]{Explained total variance in original parameters by each PC. From PC 5 on. little variance is explained by each PC, which is enlarged in Right}
    \label{fig:pc}
\end{figure}

@@

    \caption[TODO]{Fraction of the length of PCs explained by parameters}
    \label{fig:pc_pie}
\end{figure}
